---
title: "Maize Yield Analysis with Climate and Soil Data"
author: "Ãlvaro Carmona-Cabrero"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Maize Yield Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette demonstrates a workflow for preparing and analyzing maize yield data in relation to climate and soil variables. The process includes data acquisition, cleaning, and transformation, with a particular focus on incorporating climate data into yield analysis. This workflow has been adapted from a vignette of the *ERA_dev* repository.

## Setting Up the Environment

First, we set up our global settings and load the necessary packages.

```{r settings, eval=FALSE}
### Settings
PROCESS_CLIMATE_FILES <- TRUE  # Flag to process climate from ERA.Compiled

research_crop <- "Maize"
research_target <- "Crop Yield"
exclusion_buffer <- 10000
M.ORIGIN <- "1900-01-01"
cores <- max(1, parallel::detectCores() - 1) # N cores for parallel processing
ID <- "Site.Key"
```

```{r load-packages, eval=FALSE}
### Load packages
if(!require("pacman", character.only = TRUE)){
  install.packages("pacman", dependencies = T)
}
source("R/ERAWide2Long_mod.R")
source("R/ERA_convert_units.R")
source("R/map_to_usda_texture.R")
if(!require("ERAg", character.only = TRUE)){
  devtools::install_github("https://github.com/EiA2030/ERAg", dependencies = T)
}
pacman::p_load(data.table, miceadds, dplyr, ERAgON, tidyverse)
```

## Processing Climate Data

The next section processes climate data if the flag is set to TRUE. Notice that this parquet file is required to conduct the analysis "../databases/ERA/POWER.CHIRPS.parquet". It should be available in the [Zenodo record][zenodo-link].

[zenodo-link]: https://zenodo.org/records/15241882?token=eyJhbGciOiJIUzUxMiJ9.eyJpZCI6IjM0ODEwZDQzLThlYTItNDhhNC1hMDQ1LTIyMGRhMDZjYTI1NSIsImRhdGEiOnt9LCJyYW5kb20iOiJhNzk3OGE1NDQ5OTdjZjJlOGE3NzBkYmMwY2UxYzA2YiJ9.xc71ZaDofsfqKM2lYTT_8Ev227W_-_ddyB_0B3zwOxPiwIOPCHywv3XzBdu5HbNyoihWpIaUajIB2-ciCAxJjQ.

```{r climate-processing, eval=FALSE}
if (PROCESS_CLIMATE_FILES == TRUE){
  ### Load required packages for climate processing
  if(!require("pacman", character.only = TRUE)){
    install.packages("pacman", dependencies = T)
  }
  required.packages <- c("data.table", "doSNOW", "ERAg", "ERAgON", "ncdf4", 
                         "raster", "sp", "terra", "arrow", "ggplot2", "circular",
                         "zoo", "pbapply", "dismo", "miceadds")
  p_load(char=required.packages, install = T, character.only = T)
  source('R/local_CalcClimate.R')
  
  ### Create directories for outputs
  AnalysisDir <- "ERA Climate Analysis/"
  ErrorDir <- paste0(AnalysisDir, "Potential Errors/")
  CROP_dir <- paste0(AnalysisDir, "Crops/")
  
  # Create directories if they don't exist
  for(dir in c(AnalysisDir, ErrorDir, CROP_dir)) {
    if(!dir.exists(dir)) dir.create(dir)
  }
  
  ### Get ERA data and apply filters
  ERA <- ERAg::ERA.Compiled[Buffer < exclusion_buffer]
  
  ### Prepare ERA dataset
  ERA[Season.Start==1 & Season.End==1, Season:="1"
     ][Season.Start==2 & Season.End==2, Season:="2"
       ][Season.Start==1 & Season.End==2, Season:="1&2"
         ][M.Year=="", M.Year:=NA
           ][Season=="", Season:=NA]
  
  ERA[, Harvest.Start:=as.Date(Harvest.Start, "%d.%m.%Y")
     ][, Harvest.End:=as.Date(Harvest.End, "%d.%m.%Y")
       ][, Plant.Start:=as.Date(Plant.Start, "%d.%m.%Y")
         ][, Plant.End:=as.Date(Plant.End, "%d.%m.%Y")]
  
  # Add elevation data
  Physical <- data.table(ERAgON::ERA_Physical)
  ERA[!is.na(Latitude), Altitude.DEM:=Physical[match(unlist(
    ERA[!is.na(Latitude), ..ID]), unlist(Physical[, ..ID])), Altitude.mean]]
  ERA[, Altitude:=Elevation][is.na(Altitude)|Altitude=="", Altitude:=Altitude.DEM]
  
  # Subset to maize yield data
  ERA.Yields <- ERA[Out.SubInd == research_target &
                    (M.Year.Start == M.Year.End &
                     (Season %in% c(1,2) | is.na(Season))) &
                    Product %in% research_crop &
                    !grepl(" x |-", Product)]
  
  # Process POWER.CHIRPS data
  unique_site_keys <- unique(ERA.Yields$Site.Key)
  POWER.CHIRPS <- arrow::read_parquet("../databases/ERA/POWER.CHIRPS.parquet")
  POWER.CHIRPS <- POWER.CHIRPS[Site.Key %in% unique_site_keys]
  
  # Add EcoCrop data
  ERA.Yields <- cbind(ERA.Yields, ERAgON::AddEcoCrop(Products = ERA.Yields[, Product]))
  
  # Calculate ETo if absent
  if(POWER.CHIRPS[, is.null(ETo)]){
    POWER.CHIRPS[, ETo:=ERAgON::PETcalc(Tmin=Temp.Min,
                                       Tmax=Temp.Max,
                                       SRad=Solar.Rad,
                                       Wind=WindSpeed,
                                       Rain=Rain,
                                       Pressure=Pressure,
                                       Humid=Humid,
                                       YearDay=Day,
                                       Latitude=Latitude,
                                       Altitude=Altitude)[, 1]]
  }
  
  # Ensure correct data types
  POWER.CHIRPS[, Site.Key :=as.factor(Site.Key)
              ][, Year:=as.factor(Year)
                ][, Day:=as.factor(Day)
                  ][, Altitude:=as.integer(Altitude)]
  
  # Estimate missing planting dates and season lengths
  ERA.Yields <- EstPDayData(DATA=ERA.Yields)$DATA
  ERA.Yields <- EstSLenData(DATA=ERA.Yields)
  
  # Refine crop cycle lengths
  A <- ERA.Yields
  A <- unique(A[!(is.na(Plant.Start)|is.na(Plant.End)|is.na(Harvest.End)|is.na(Harvest.End))
              ][, PLANT.AVG:=Plant.Start+(Plant.End-Plant.Start)/2
                ][, HARVEST.AVG:=Harvest.Start+(Harvest.End-Harvest.Start)/2
                  ][!(((Plant.End-Plant.Start)>7)|((Harvest.End-Harvest.Start)>7))
                    ][, SLEN:=round(HARVEST.AVG-PLANT.AVG,0)
                      ][!SLEN<30, c("Product", ..ID, "Code", "Variety", "M.Year", "PLANT.AVG", "SLEN", "Country")])
  
  B <- A[, list(SLEN.mean=mean(SLEN), SLEN.med=median(SLEN), N=.N), by=c("Product")
       ][, SLEN.mean:=round(as.numeric(SLEN.mean), 0)
         ][, SLEN.med:=round(as.numeric(SLEN.med), 0)]
  
  B[, SLEN.EcoC:=round(apply(ERA.Yields[match(B$Product, ERA.Yields$Product), c("cycle_min","cycle_max")], 1, mean), 0)]
  
  # Use estimated crop lengths where we have enough observations
  B <- B[N>=5]
  fwrite(B, paste0(CROP_dir, "Crop Season Lengths Estimated From Data.csv"))
  
  # Update EcoCrop values with our estimates
  N <- match(ERA.Yields$Product, B$Product)
  N1 <- which(!is.na(N))
  ERA.Yields$cycle_min[N1] <- B$SLEN.med[N[N1]]
  ERA.Yields$cycle_max[N1] <- B$SLEN.med[N[N1]]
  
  # Refine planting dates using rainfall data
  ERA.Yields <- EstPDayRain(Data=ERA.Yields, 
                           ID=ID, 
                           Rain.Data=POWER.CHIRPS,
                           Rain.Data.Name="CHIRPS", 
                           Rain.Field="Rain", 
                           DaysBefore=2,
                           MultiplyWin=1, 
                           Window=c(14,14), 
                           Widths=c(2,3,3), 
                           Rain.Thresholds=c(30,20,15), 
                           Uncertainty.Min=7, 
                           Uncertainty.Max=90, 
                           Add.Values=T, 
                           Use.Data.Dates=T)
  
  # Calculate climate variables
  Climate <- local_CalcClimate(DATA=ERA.Yields,
                             ID=ID,
                             CLIMATE=POWER.CHIRPS,
                             Rain.Data.Name="CHIRPS", 
                             Temp.Data.Name="POWER",
                             Rain.Windows=c(6*7,4*7,2*7,2*7),
                             Widths=c(3,3,2,2),
                             Rain.Threshold=c(30,30,20,15), 
                             Win.Start=-3,
                             Max.LT.Avg=2010, 
                             Do.LT.Avg=T, 
                             Do.BioClim=T, 
                             Windows=data.table(Name="Plant.1-30", Start=1, End=30), 
                             SaveDir=AnalysisDir,
                             ErrorDir=NA,
                             ROUND=3)
}
```

## Data Preparation for Analysis

After processing the climate data, we proceed with loading the processed data, preparing it for analysis, and merging climate, soil, and yield data.

```{r load-processed-data, eval=FALSE}
### Change ID column for merging
ID <- "ID"

### Set working directory for climate data
ClimDir <- "ERA Climate Analysis/Analysis/422814143322303020152010130-3POWCHIFALSE0.6/"
if(!dir.exists(ClimDir)){
  dir.create(ClimDir)
}

### Load processed data
ClimateA <- load(paste0(ClimDir, "ClimStatsA.RData"))
ERA <- load(paste0(ClimDir, "Data.RData"))
Soils <- ERAgON::ERA_ISDA

### Prepare ERA data and apply filters
setDT(DATA)
DATA <- data.table::copy(DATA)

climate_ids <- unique(Seasonal$Observed[, ID])

DATA <- DATA[Product == research_crop &
             !grepl("\\.\\.", Site.ID) &
             Irrigation.C == FALSE & Irrigation.T == FALSE &
             Site.Type %in% c("Station", "Station.Farm", "Farm", "Farm & Station") &
             (is.na(Mulch.Code) | Mulch.Code != "b26") &
             Code != "EO0006" &
             Site.Key %in% climate_ids &
             nchar(M.Year) < 7 &
             !(T1 == "b26" | T2 == "b26" | T3 == "b26" | T4 == "b26" | 
               T5 == "b26" | T6 == "b26" | T7 == "b26" | T8 == "b26")]

era_ids <- unique(DATA[, ID])

### Basic data check
high_yields <- DATA[MeanT > 10000, ]
print(DATA[Code %in% high_yields$Code, MeanT])
hist(DATA[, MeanT])
```

## Preparing Soil Data

Next, we prepare the soil data for integration with the climate and yield data.

```{r prepare-soils, eval=FALSE}
### Prepare Soils data
Soils <- Soils[Site.Key %in% era_ids, ]

# Process depth information
Soils[grepl("0..200cm", Variable), Depth.Upper:=0
     ][grepl("0..200cm", Variable), Depth.Lower:=200
       ][, Variable:=gsub("_m_30m_0..200cm", "", Variable)]

Soils[grepl("0..20cm", Variable), Depth.Upper:=0
     ][grepl("0..20cm", Variable), Depth.Lower:=20
       ][, Variable:=gsub("_m_30m_0..20cm", "", Variable)]

Soils[grepl("20..50cm", Variable), Depth.Upper:=20
     ][grepl("20..50cm", Variable), Depth.Lower:=50
       ][, Variable:=gsub("_m_30m_20..50cm", "", Variable)]

# Calculate weighted averages by depth
Soils[, Weight:=Depth.Lower-Depth.Upper]

Soils <- unique(Soils[, list(Median=weighted.mean(Median, Weight),
                           Mean=weighted.mean(Mean, Weight),
                           SD=weighted.mean(SD, Weight),
                           Mode=Mode[Weight==max(Weight)]),
                    by=list(Variable, Site.Key)])

# Convert to wide format
Soils <- dcast(Soils, Site.Key~Variable, value.var=c("Median", "Mean", "SD", "Mode"))
```

## Merging Climate and Soil Data with ERA

Here we merge the climate and soil data with the ERA data to create integrated datasets.

```{r merge-data, eval=FALSE}
### Prepare Long Term and Observed Climate data
Clim.Data <- Seasonal[["Observed"]]
Clim.Data <- Clim.Data[ID %in% era_ids, ]

# Create merge keys
Clim.Data[, MCode:=paste(ID, EU, PD.Used, sep="|")]
DATA[, MCode:=paste(Site.Key, EU, P.Date.Merge, sep="|")]

LTAvg.Clim.Data <- Seasonal[["LongTerm"]][["LT.Clim.Avg"]]
LTAvg.Clim.Data <- LTAvg.Clim.Data[ID %in% era_ids]

# Create long-term climate merge keys
LTAvg.Clim.Data[, MCode.LT:=paste(ID, EU, Season)]
Clim.Data[, MCode.LT:=paste(ID, EU, Season)]

### Merge Climate & Soil Data with ERA for each window
ERA.Clim.Soil <- lapply(Clim.Data[, unique(W.Name)], FUN=function(X){
  
  # Subset seasonal climate data to calculation window X
  CData <- Clim.Data[W.Name==X]
  
  # Subset long-term climate averages
  LT.Clim.Data <- LTAvg.Clim.Data[W.Name==X]
  LT.Clim.Data.Mean <- LT.Clim.Data[Variable=="Mean"]
  LT.Clim.Data.Median <- LT.Clim.Data[Variable=="Median"]
  
  # Select columns for deviation calculation
  Cols <- colnames(CData[, !c("Tmax.sd", "Tmean.sd", "PD.Used", "W.Start", "W.End", 
                             "W.Name", "ID", "M.Year", "Season", "MCode", "MCode.LT", 
                             "EU", "ETo.NA")])
  
  # Calculate deviations from long-term averages
  Mean.Diff <- CData[, ..Cols] - 
    LT.Clim.Data.Mean[match(CData[, MCode.LT], LT.Clim.Data.Mean[, MCode.LT])][, ..Cols]
  Median.Diff <- CData[, ..Cols] - 
    LT.Clim.Data.Median[match(CData[, MCode.LT], LT.Clim.Data.Median[, MCode.LT])][, ..Cols]
  
  # Rename deviation columns
  colnames(Mean.Diff) <- paste0(colnames(Mean.Diff), ".Dev.Mean")
  colnames(Median.Diff) <- paste0(colnames(Median.Diff), ".Dev.Median")
  
  # Join data
  CData <- cbind(CData, Mean.Diff, Median.Diff)
  SoilData <- Soils[match(DATA[, Site.Key], Site.Key)]
  
  # Combine all data
  cbind(DATA, 
        CData[match(DATA$MCode, MCode), !c("MCode", "EU", "ID")], 
        SoilData[, !"Site.Key"])
})

# Name list elements
names(ERA.Clim.Soil) <- unique(Clim.Data$W.Name)
```

## Creating Long Format Datasets

Convert the wide format datasets to long format for each estimation method.
Four climate datasets were produced based on the calculation period:
- data_ERA.Clim: Only use observations that report planting and harvest date. Climate factors are computed using these dates.
- EcoCrop_ERA.Clim: Use reported planting and harvest date. Estimate dates for the rest of the observations (see Methods section). Climate factors are computed using these dates.
- Plant1_30_ERA.Clim: Use reported planting and harvest date. Estimate dates for the rest of the observations (see Methods section). Climate factors are computed from planting date to 30 days after planting. 
- PreSowing_ERA.Clim: Use reported planting and harvest date. Estimate dates for the rest of the observations (see Methods section). Climate factors are computed from 30 days before planting to planting date.

```{r create-long-format, eval=FALSE}
# Convert to long format
data_ERA.Clim.Soil <- ERAWide2Long_mod(ERA.Clim.Soil$Data)
EcoCrop_ERA.Clim.Soil <- ERAWide2Long_mod(ERA.Clim.Soil$EcoCrop)
Plant1_30_ERA.Clim.Soil <- ERAWide2Long_mod(ERA.Clim.Soil$`Plant.1-30`)
PreSowing_ERA.Clim.Soil <- ERAWide2Long_mod(ERA.Clim.Soil$PreSowing)

# Remove duplicate columns
unique_cols <- !duplicated(names(data_ERA.Clim.Soil))
data_ERA.Clim.Soil <- data_ERA.Clim.Soil[, ..unique_cols]

unique_cols <- !duplicated(names(EcoCrop_ERA.Clim.Soil))
EcoCrop_ERA.Clim.Soil <- EcoCrop_ERA.Clim.Soil[, ..unique_cols]

unique_cols <- !duplicated(names(Plant1_30_ERA.Clim.Soil))
Plant1_30_ERA.Clim.Soil <- Plant1_30_ERA.Clim.Soil[, ..unique_cols]

unique_cols <- !duplicated(names(PreSowing_ERA.Clim.Soil))
PreSowing_ERA.Clim.Soil <- PreSowing_ERA.Clim.Soil[, ..unique_cols]

# Remove duplicates
data_ERA.Clim.Soil <- unique(data_ERA.Clim.Soil)
EcoCrop_ERA.Clim.Soil <- unique(EcoCrop_ERA.Clim.Soil)
Plant1_30_ERA.Clim.Soil <- unique(Plant1_30_ERA.Clim.Soil)
PreSowing_ERA.Clim.Soil <- unique(PreSowing_ERA.Clim.Soil)

# Rename column in presowing dataset
setnames(PreSowing_ERA.Clim.Soil, "Rain.sum", "Presowing.Rain.sum")
```

## Final Data Processing and Export

Process each dataset to handle NA values, convert units, and filter outliers before exporting.

The presowing dataset (PreSowing_ERA.Clim) is modified to include all climate factors in EcoCrop_ERA.Clim.Soil and the presowing total rainfall is added.

```{r final-processing, eval=FALSE}
### Process and export each dataset
datasets = list(data_ERA.Clim.Soil, EcoCrop_ERA.Clim.Soil, 
                Plant1_30_ERA.Clim.Soil, PreSowing_ERA.Clim.Soil)
datasets_names = list("data_ERA_Clim_Soil", "EcoCrop_ERA_Clim_Soil", 
                      "Plant1_30_ERA_Clim_Soil", "PreSowing_ERA_Clim_Soil")

for (i in 1:length(datasets)){
  dataset <- datasets[[i]]
  dataset_name <- datasets_names[[i]]
  dataset <- dataset[, .SD, .SDcols = unique(names(dataset))]
  
  # Add presowing rainfall to other datasets
  if (dataset_name != "PreSowing_ERA_Clim_Soil"){
    setkey(dataset, Site.Key, TID, Code, T.Descrip, MeanT)
    setkey(PreSowing_ERA.Clim.Soil, Site.Key, TID, Code, T.Descrip, MeanT)
    copy <- copy(dataset)
    copy[PreSowing_ERA.Clim.Soil, Presowing.Rain.sum := i.Presowing.Rain.sum]
    dataset <- copy
    predictors <- colnames(dataset)[c(128:160, 161:210, 211:283)]
    predictors <- append(predictors, colnames(dataset)[length(colnames(dataset))])  
  } else{
    predictors <- colnames(dataset)[c(128:160, 161:210, 211:283)]
  }
  
  # Remove rows with NA in predictors
  dataset_noNA <- dataset[!is.na(get(predictors[1])) & 
                           rowSums(is.na(dataset[, ..predictors])) == 0]
  
  # Convert to data.frame and add USDA soil texture
  dataset_noNA_df <- as.data.frame(dataset_noNA)
  dataset_noNA_df <- dataset_noNA_df %>%
    mutate(Soil.Texture.USDA = sapply(Soil.Texture, map_to_usda_texture))
  
  # Convert units
  dataset_noNA_df <- ERA_convert_units(dataset_noNA_df)
  
  # Create histogram
  hist(dataset_noNA_df$MeanT, main=dataset_name)
  
  # Check low yields
  low_yields <- dataset_noNA_df %>% filter(MeanT<100) %>% dplyr::select(MeanT)
  hist(low_yields$MeanT, main=paste(dataset_name, "Low yields"))
  
  # Apply filters (except for presowing dataset)
  if (dataset_name != "PreSowing_ERA_Clim_Soil"){
    dataset_noNA_df <- dataset_noNA_df %>%
      filter(!MeanT < 100,
             Rain.sum > 100)
  }
  
  # Rename yield column
  dataset_noNA_df <- dataset_noNA_df %>%
    rename(yield = MeanT)
  
  # Export datasets
  write_csv(dataset_noNA_df, paste0("outputs/era/psw_", dataset_name, ".csv"))
  saveRDS(dataset_noNA_df, file=paste0("outputs/era/psw_", dataset_name, ".rds"))
}
```

## Additional Data Checks

Perform a few additional checks on yields to identify potential issues.

```{r data-checks, eval=FALSE}
### Additional data quality checks
count(dataset_noNA_df %>% filter(yield < 1000) %>% dplyr::select(yield))

# Check experiments with low yields
low_yield_codes <- unique(dataset_noNA_df %>% filter(yield < 1000) %>% dplyr::select(Code))

low_yield_exp <- dataset_noNA_df %>%
  filter(Code %in% low_yield_codes$Code) %>%
  dplyr::select(Code, yield, TorC)

# Check consistently low yield experiments (may be due to units)
consistent_low_yield_exp <- dataset_noNA_df %>%
  group_by(Code) %>%
  filter(all(yield < 1000)) %>%
  ungroup() %>%
  dplyr::select(Code, yield, TorC, OriUnits)

# Check specific experiment
DATA %>% filter(Code == "NN0297") %>% dplyr::select(MeanT, MeanC, Code, Units)
```

## Conclusion

This vignette demonstrates a comprehensive workflow for processing and analyzing maize yield data in relation to climate and soil factors. The resulting datasets can be used for further analysis, modeling, and visualization to understand the impacts of various environmental factors on maize yields.
