---
title: "Comparing SoilGrids and ERA Soil Properties"
author: "Álvaro Carmona-Cabrero"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 8
    fig_height: 6
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 6,
  out.width = "100%"
)
```

## Introduction

This vignette demonstrates how to compare soil properties between two data sources: SoilGrids and ERA databases. We'll examine correlations between key soil properties (organic carbon, pH, clay, sand, and silt) and visualize the relationships using scatter plots and correlation matrices.

## Setup and Data Loading

First, we load the necessary libraries and our dataset:

```{r load-libraries-data}
# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(corrplot)
library(data.table)

# Load dataset
data <- read.csv("outputs/era/base_ecocrop.csv")
data <- unique(data)

# pH correction (dividing by 10)
data$Mean_ph_h2o <- data$Mean_ph_h2o/10

# Preview data structure
str(data[, c("Mean_oc", "SOC", "Mean_ph_h2o", "Soil.pH", "Mean_clay_tot_psa", 
            "CLY", "Mean_sand_tot_psa", "SND", "Mean_silt_tot_psa", "SLT")])
```

## Data Preparation

Let's rename variables for clarity and consistency:

```{r rename-variables}
data_renamed <- data %>%
  rename(
    # SoilGrids variables
    soilgrids_oc = Mean_oc,
    soilgrids_ph = Mean_ph_h2o,
    soilgrids_clay = Mean_clay_tot_psa,
    soilgrids_sand = Mean_sand_tot_psa,
    soilgrids_silt = Mean_silt_tot_psa,
    
    # ERA variables
    era_oc = SOC,
    era_ph = Soil.pH,
    era_clay = CLY,
    era_sand = SND,
    era_silt = SLT
  )

# Preview the renamed data
head(data_renamed[, c("soilgrids_oc", "era_oc", "soilgrids_ph", "era_ph", 
                     "soilgrids_clay", "era_clay", "soilgrids_sand", "era_sand", 
                     "soilgrids_silt", "era_silt")])
```

## Create Comparison Function

Define a function to create consistent scatter plots for comparing soil properties:

```{r comparison-function}
# Function to create cleaner comparison plots for soil properties
compare_soil_properties <- function(data, soilgrids_var, era_var, x_label, y_label) {
  complete_data <- data %>% 
    filter(!is.na(!!sym(soilgrids_var)), !is.na(!!sym(era_var)))
  
  pearson_corr <- cor(complete_data[[soilgrids_var]], complete_data[[era_var]], 
                      use = "complete.obs", method = "pearson")
  
  p <- ggplot(complete_data, aes_string(x = soilgrids_var, y = era_var)) +
    geom_point(alpha = 0.5) +
    geom_smooth(method = "lm", color = "red") +
    labs(
      subtitle = paste0("r = ", round(pearson_corr, 3)),
      x = x_label,
      y = y_label
    ) +
    theme_minimal() +
    theme(plot.title = element_blank(),
          plot.background = element_rect(fill = "white", color = NA),
          panel.background = element_rect(fill = "white", color = NA),
          panel.grid.major = element_line(color = "gray90"),
          panel.grid.minor = element_line(color = "gray95"))
  
  return(p)
}
```

## Define Soil Property Comparisons

Specify the soil properties to compare:

```{r define-comparisons}
# Define soil property pairs to compare with clean labels
soil_comparisons <- list(
  list(soilgrids_var = "soilgrids_oc", era_var = "era_oc", 
       x_label = "SoilGrids OC", y_label = "ERA OC"),
  list(soilgrids_var = "soilgrids_ph", era_var = "era_ph", 
       x_label = "SoilGrids pH", y_label = "ERA pH"),
  list(soilgrids_var = "soilgrids_clay", era_var = "era_clay", 
       x_label = "SoilGrids Clay", y_label = "ERA Clay"),
  list(soilgrids_var = "soilgrids_sand", era_var = "era_sand", 
       x_label = "SoilGrids Sand", y_label = "ERA Sand"),
  list(soilgrids_var = "soilgrids_silt", era_var = "era_silt", 
       x_label = "SoilGrids Silt", y_label = "ERA Silt")
)
```

## Generate and Display Comparison Plots

Create scatter plots for each soil property comparison:

```{r generate-plots, fig.width=8, fig.height=6}
# Generate plots
scatter_plots <- list()
for (comp in soil_comparisons) {
  scatter_plots[[comp$x_label]] <- compare_soil_properties(
    data_renamed, comp$soilgrids_var, comp$era_var, comp$x_label, comp$y_label)
  print(scatter_plots[[comp$x_label]])
  ggsave(paste0("outputs/era/soil_comparison_", comp$era_var, ".png"))
}
```

## Correlation Analysis

Calculate and visualize the correlation matrix for all soil variables:

```{r correlation-matrix}
soil_vars <- c("soilgrids_oc", "era_oc", "soilgrids_ph", "era_ph", 
               "soilgrids_clay", "era_clay", "soilgrids_sand", "era_sand", 
               "soilgrids_silt", "era_silt")

# Filter complete cases for correlation analysis
soil_data <- data_renamed %>% select(all_of(soil_vars)) %>% drop_na()

# Calculate Pearson correlation
cor_matrix <- cor(soil_data, method = "pearson")

# Define clean labels for the correlation plot
clean_labels <- c("SG OC", "ERA OC", "SG pH", "ERA pH", 
                  "SG Clay", "ERA Clay", "SG Sand", "ERA Sand", 
                  "SG Silt", "ERA Silt")

# Rename matrix dimensions
colnames(cor_matrix) <- clean_labels
rownames(cor_matrix) <- clean_labels

# Display correlation matrix
round(cor_matrix, 2)
```

## Correlation Visualization

Create a correlation plot to visualize the relationships:

```{r correlation-plot, fig.height=8, fig.width=8}
# Create correlation plot
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45,
         tl.cex = 0.8, 
         diag = FALSE,
         title = "Correlation Between SoilGrids and ERA Soil Properties",
         mar = c(0,0,1,0))

# Save correlation plot
png("outputs/era/soil_correlation_plot.png", width = 8, height = 8, units = "in", res = 300)
corrplot(cor_matrix, method = "color", type = "upper", 
         tl.col = "black", tl.srt = 45,
         tl.cex = 0.8,
         diag = FALSE)
dev.off()
```
